\emph{icepack}: a novel glacier flow modeling package

### 2020-11-25 icepack.tex
Add ice sheet figure

As a second test case for icepack, we ran an experiment inspired by \cite{kessler2008fjord}.
In the original work, the authors couple an ice-flow model with simple models for bed-erosion, calving, and \textcolor{red}{glacial isostatic adjustment (GIA)} to investigate the formation of deep fjords that are characteristic of coastlines on Baffin Island, Greenland, and British Columbia, among others.
Their first experiment is set in an idealized circular domain that contains a bedrock plateau rimmed by a ridge that is studded with four valleys.
They find that fjords become well-developed after about one million years and that ice velocities are highest in these fjords.

Similar to \cite{kessler2008fjord}, we use the shallow ice model; however, we do not couple it to bed-erosion, calving, or GIA.
We initialized the experiment with a simple but unrealistic ice thickness \textcolor{red}{(figure X)}.
We then evolve the ice sheet without climate forcing for 500 years until it is near equilibrium.
The time it takes to reach equilibrium is much shorter than the \cite{kessler2008fjord} experiment because the bedrock is static.
The first part of our simulation is an exponential relaxation towards equilibrium.
A velocity pattern emerges that is similar to results from \cite{kessler2008fjord}, with the highest velocities where ice is constricted through the deepest valleys \textcolor{red}{(figure X)}.
In and upstream of the valleys, the surface is drawn down due to the elevated export of ice \textcolor{red}{as shown in figure X}.

The ice-flow model chosen for this simulation is computationally cheap compared to the other ice-flow models, such that it can be used to simulate an ice sheet over thousands of years.
By simply switching between ice-flow models, icepack is able to simulate many different glaciological conditions over a wide range of spatiotemporal scales.
For example, icepack can be used to simulate a small domain dominated by an ice stream, an ice catchment that includes slow-moving interior ice and a floating ice shelf, or, as in this example, an entire ice sheet.

>>>>

\begin{figure}[h]
    \includegraphics[width=0.95\linewidth]{demos/ice-sheet/ice-sheet.png}
    \caption{Synthetic ice sheet simulation: (a) the bed elevation profile consists of a ring of mountains with valleys of varying depth surrounding a flat plateau where the ice is initialized to (b) the change in thickness over the simulation period}
    \label{fig:ice-sheet}
\end{figure}

As a demonstration of the shallow ice approximation model, we ran an experiment inspired by \cite{kessler2008fjord}.
In this work, the authors coupled an ice flow model to simple models for bed-erosion, calving, and glacial isostatic adjustment (GIA) to investigate the formation of deep fjords that are characteristic of coastlines on Baffin Island, Greenland, and British Columbia, among others.
We emulated their domain and bedrock geometry -- a plateau surrounded by a ridge punctuated by four valleys -- without also simulating erosion or GIA.
These effects will be the subject of future work.

Figure \ref{fig:ice-sheet} shows the results of this computational experiment.
We initialized the experiment with a simple but unrealistic ice thickness.
We then evolved the ice sheet without climate forcing for 500 years.
The ice sheet relaxes very rapidly in the first 200 years, but after this changes are much slower as ice must be funneled through one of the four narrow valleys.
The resulting velocity pattern is similar to that of \citet{kessler2008fjord}, with the highest velocities where ice flows out from the valleys.
In and upstream of the valleys, the surface is drawn down due to the elevated export of ice.

The diagnostic model used in this demonstration is cheap enough that it can be used to simulate ice sheets over several millenia in a matter of minutes on a desktop.
A key feature of icepack is the ability to choose between many different diagnostic models.
Users are free to decide whichever model is most appropriate for the spatial and temporal scales of their problem.
